---
title: "Pi Day Creation"
author: "Chris van Hasselt"
date: "March 14, 2016"
output: html_document
---

```{r setup, echo=FALSE}

# load libraries
library(plotrix)
library(RColorBrewer)
```

```{r loadData,echo=FALSE}
# Grab raw digits from file (first 10K digits of fractional pi)
rawPiDigits <- readLines("PI10K_DP.txt")

# convert to a data.frame
piDigits <- data.frame(digit = as.numeric(unlist(strsplit(rawPiDigits,""))))

# create row index for the list of digits (I'd like a better solution to this)
piDigits$idx <- 1:nrow(piDigits) -1

# compute remainder, mod 360 for the degree measure
piDigits$degs<- sapply(piDigits$idx,FUN=function(x){
    as.numeric(x %% 360)
})

# quotient mod 360
piDigits$q <- sapply(piDigits$idx,FUN=function(x){
    as.numeric((x %/% 360) + 1) 
})

# normalize by the maximum rotation of degrees
norm <- max(piDigits$q)

# create color mapping function and map 10 colors
piColorFunc <- colorRampPalette(brewer.pal(10,"Paired"))
piColors <- piColorFunc(10)


```


```{r plotting, echo=FALSE, fig.width=8, fig.height=8, fig.align='center'}
# create plot, add text
polar.plot((piDigits$digit * piDigits$q)/norm,piDigits$degs,
           show.grid=FALSE,
           line.col=piColors[piDigits$digit + 1],
           lwd=.5,do.first=plot_bg(col="black"))
text(0,0, expression(pi),cex=8,adj=.5,col='black')
text(0,0, expression(pi),cex=8,adj=c(.52,.52),col=piColors[1])
```

## About the Visualization

This visualization of pi does not demonstrate any deep principles about pi.  Pi
is an irrational numbeer, the ratio of a circle's radius to it's perimeter. The 
digits of pi form an apparently random sequence of numbers. However, despite 
it's random apparently nature it is fundamental to many equations in mathematics.

This visualization takes the fractional portion of pi and arranges the digits in sequence
around a circle, mapping each digit to its position modulo 360 in the sequence.
The whole number quotient of the modular division is used as a scaling factor 
for each digit, and each digit is mapped to a different color.  The
first 10,000 digits of pi's fractional portion are plotted.

The visualization uses the the _plotrix_ R library for plotting, and the 
_RColorBrewer_ library for mapping colors.  The code for data manipulation is below.

```{r data_code, ref.label='loadData', echo=TRUE}

```

The code for creating the plot is below.

```{r plotting_code, ref.label='plotting', echo=TRUE, eval=FALSE}

```




